---
title: "Bike-share Case Study"
author: "Alexander McDaniel"
date: "2024-01-31"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{=html}
<style type="text/css">

h1.title {
  font-size: 38px;
  color: Black;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: Dark Blue;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: Dark Blue;
  text-align: center;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Capstone Project

### Background Scenario

For this Data Analysis Project we will be working with a fictional
company, "Cyclistic" a bike-sharing company. The scenario will have
myself as a junior data analyst working with the company marketing
analyst team to help understand the difference between the casual riders
vs. the annual memberships.Gathering insights, I will design a new
marketing strategy to convert casual riders to annual members. Using
compelling data insight and professional data visualizations I will
receive approval from the Cyclistic executives to act on my
recommendations.

### Stakeholders

-   Director of Marketing ("Lily Moreno")
-   Cyclistic Executives Team
-   Cyclistic Marketing analytics team

## Ask

### Business Objective

#### Questions:

1.  How do annual members and casual riders sue Cyclistic bikes
    differently, and is it quantifiable?
2.  What hypothesis can we generate that could be the reason for the
    differences
3.  Is there executable measures the company can make to convert more
    casual riders to have a annual membership

## Preparing the Data

#### Importing library

```{r import Library}
install.packages("tidyverse")
install.packages("readr")
install.packages("plyr")
install.packages("purrr")
library("purrr")
library("tidyverse")
library("ggplot2")
library("dplyr")
library("readr")
library("plyr")

```

### Importing Data-sets

The Cyclistic has has provided records of both casual and annual
memberships trip data-sets in from the year 2023. Due to the company
providing the data, we can consider the data first party and credibly.
The data is was made available by Motivate International Inc. under the
[license](https://divvybikes.com/data-license-agreement).

#### Loading .csv file

```{r Importing all DataSets}
##Jan2023 <- read_csv("202301-divvy-tripdata.csv")
##Feb2023 <- read_csv("202302-divvy-tripdata.csv")
##Mar2023 <- read_csv("202303-divvy-tripdata.csv")
##Apr2023 <- read_csv("202304-divvy-tripdata.csv")
##May2023 <- read_csv("202305-divvy-tripdata.csv")
##Jun2023 <- read_csv("202306-divvy-tripdata.csv")
##Jul2023 <- read_csv("202307-divvy-tripdata.csv")
##Aug2023 <- read_csv("202308-divvy-tripdata.csv")
##Sep2023 <- read_csv("202309-divvy-tripdata.csv")
##Oct2023 <- read_csv("202310-divvy-tripdata.csv")
##Nov2023 <- read_csv("202311-divvy-tripdata.csv")
##Dec2023 <- read_csv("202312-divvy-tripdata.csv")
```

#### ROCCC Data Check

Now that I received the data I need to check for any potential bias in
the data provided. using the ROCCC Method(reliable, original,
comprehensive, current and cited) I examine the datasets keeping the
data Check in mind.

## Processing the Data

Sample Table

```{r Test Run of DataSet}
## Jan2023 <- read_csv("202301-divvy-tripdata.csv")
## head(Jan2023)
```

### Observations:

From Looking over the data we can see that we have 13
columns(variables). analyzing the data we will find the data is
complete, missing information can be seen in the data-set such at
Starting_station_name, start_station_id, end_station_name, and
end_station_id. Another issue that can be observed regarding the
stations ID is that the format is not a universal format. When thinking
to the Business objective we must solve, the follow data that is missing
wont affect the analyzing process.

### Calculating for Confidence level and Margin error:

Calculated as Followed:

Population size of Chicago: 8,927,000

Sample size: 5,719,877

Confidence level:

Margin of Error:

### Combining all the Data-sets together

```{r Creating dataset to contain all data}
# setwd("E:/Data Analytics/Datasets/Cyclistics dataset")
# getwd()
# 
# Data_2024 <- list.files(                 ## Creating a dataset to combine all .csv
#   pattern = "*.csv",
#   full.names = T) %>%
#   lapply(read_csv) %>%
#   bind_rows()
# 
# write.csv(Data_2024, "Data_2024.csv")    ## Exporting Data_2024 to save Environment 
# 
# head(Data_2024)                          ## Sample of Table
```

### Cleaning Data-set

Thoughts: First thing that come to mind is I need to change the
starting/end time to values that I can graph. Finding the difference
between the start and end will give me the duration of the trip.I will
also need to calculate the distance that was traveled for the trip, we
do not have a tracker for distance so the distance between the start
station and end station will be a sufficient alternative.Lastly, Finding
the days of the week that the difference members chose to ride and the
time will be important.

Predicted Visualizations needed:

-   1\. Graph of Starting / end stations

-   2\. Days of week cyclist ride.

-   3\. Times rider during the day

-   4\. Distances between the memberships they ride.

First Lets create the new column needed to organize the data, The Create
a new data with the new variables we will need.

```{r create data set}
### Create a new data set, trim the variable that are not needed
Day_of_Week <- Data_2024                                   # Replicate data of Data_2024
Day_of_Week <- Day_of_Week[ , -c(1:3, 6:9)]               # Deleted the Variables not needed

### Convert date's to days of the week.
Day_of_Week$weekday <- weekdays(Day_of_Week$started_at)    # Convert Date to Days of the week
head(Day_of_Week)
```

```{Add durations varaibles}
Day_of_Week$Duration <- difftime(Day_of_Week$ended_at,Day_of_Week$started_at, units = "mins")
```

```{r Separate data from times}
#  separate the times of day to separate column
# Day_of_week  <- separate(Day_of_week,
#          col = ended_at,
#          into = c("ended_date", "ended_time"),
#          sep = " ",
#          remove = FALSE)
# 
# Day_of_week  <- separate(Day_of_week,
#          col = started_at,
#          into = c("Start_date", "Start_time"),
#          sep = " ",
#          remove = FALSE)


```

```{R Removed Chuck code}
### Duration of each trip
class(Day_of_week$started_at)



# Day_of_week$ <- df %>%
#   mutate(time_diff = as.character(hms::as_hms(
#     difftime(
#       time1 = ended_at,
#       time2 = started_at,
#       units = "mins"
#     )
#   )))
# 
# 
# 
# Day_of_week$Start_time <- hms(Day_of_week$Start_time)
# Day_of_week$Start_time <- as.numeric(Day_of_week$Start_time)
# 
# Day_of_week$ended_time <- hms(Day_of_week$ended_time)
# Day_of_week$ended_time <- as.numeric(Day_of_week$ended_time)

```

## Analyzing Data

```{r starting hours of day}
Day_of_Week %>%
  mutate(hour = hour(started_at)) %>%
  group_by(hour) %>%
  tally() %>%
  ggplot() +
  geom_line(aes(x = hour, y = n)) +
  scale_x_continuous(breaks = seq(0, 24, 4))


```

## Sharing the Data

## Act
